---
title: "Pair Correlation Analysis"
author: "syslo"
date: "2023-01-26"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Overview
  The purpose of this code is to generate a graph depicting the pair correlation function of bcells with other cells in the tumor environment
  
## Loading Packages
```{r load-pkgs}
# install.packages("raster")
# install.packages("plyr")
# install.packages("terra")

library(raster)
library(terra)
library(plyr)
library(data.table)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## File Setup
File system is setup as follows:
Main_directory
  - PCF.Rproj (this R project)
  - data (directory containing multiplex data files)
  - reference (directory containing "conditions.csv", similar to sample sheet)
  - results (output directory for saving results)
  - source_code (directory containing function scripts)
  
```{r}
# Set working directory to main directory
data_loc <- "./data"

sample_id <- list.files(path = data_loc, pattern = "*.txt", full.names = TRUE)
conditions <- as.data.frame(sample_id) %>%
  mutate(group = "")
write.csv(conditions, file = "./reference/conditions.csv")

# Open conditions.csv in Excel and define sample groups
# For this analysis, we do not define separate sample groups so all samples are defined with group "base"

# RUN THE CODE BELOW EACH TIME YOU RE-RUN PCF ANALYSIS
conditions <- read.csv("./reference/conditions.csv") %>%
  dplyr::select(sample_id, group)
groups <- unique(conditions$group)
```

# PCF Analysis
The following are the code for the pcf analyses. The key for each analysis and functions file is as follows:
•	A = All tumor samples using B cells as the focus corresponding to nearby TAMs, Tumor cells, T cells, and Microglia
•	B = All tumor samples using B cells as the focus corresponding to nearby TGFB+ TAMs, Tumor cells, T cells, and Microglia
•	C = All tumor samples using B cells as the focus corresponding to nearby TGFB+ TAMs and Tumor cells

## PCF Analysis A
This analysis will contain all samples. We use B cells as focus point, comparing to TAMs, tumor cells, t cells, and microglia.
```{r pcf_a}
# Load packages
library(doParallel)
library(parallel)
library(foreach)

# Make sure this is the right functions file
source(file = "./source_code/01_functions_a.R")

no_cores <- detectCores()
registerDoParallel(makeCluster(no_cores))


for (i in groups) {
  df <- paste(i, "df", sep = "_")
  
  ls <- conditions %>%
    filter(group == i) %>%
    select_('sample_id')
  
  assign(df,
         foreach(id = 1:nrow(ls)) %dopar% read_mpx(ls, id))

  df <- get(df)

  # TAM PCF
  pcf1 <- paste(i, "pcf_tam", sep="_")

  assign(pcf1,
       foreach(id = 1:length(df)) %dopar% pcf_tam(df[[id]]))
  
  pcf1_all <- get(pcf1)
  pcf1_sum <- head(pcf1_all[[1]],n=200)
  for (j in 2:length(pcf1_all)){
    pcf1_sum <- pcf1_sum + head(pcf1_all[[j]],n=200)
  }
  pcf1_avg <- pcf1_sum/length(pcf1_all)
  pcf1_avg <- pcf1_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf1, pcf1_avg)
  
  
  # TUMOR PCF
  pcf2 <- paste(i, "pcf_tumor", sep="_")

  assign(pcf2,
       foreach(id = 1:length(df)) %dopar% pcf_tumor(df[[id]]))
  
  pcf2_all <- get(pcf2)
  pcf2_sum <- head(pcf2_all[[1]],n=200)
  for (j in 2:length(pcf2_all)){
    pcf2_sum <- pcf2_sum + head(pcf2_all[[j]],n=200)
  }
  pcf2_avg <- pcf2_sum/length(pcf2_all)
  pcf2_avg <- pcf2_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf2, pcf2_avg)

  # T_CELL PCF
  pcf3 <- paste(i, "pcf_tcell", sep="_")

  assign(pcf3,
       foreach(id = 1:length(df)) %dopar% pcf_tcell(df[[id]]))

  pcf3_all <- get(pcf3)
  pcf3_sum <- head(pcf3_all[[1]],n=200)
  for (j in 2:length(pcf3_all)){
    pcf3_sum <- pcf3_sum + head(pcf3_all[[j]],n=200)
  }
  pcf3_avg <- pcf3_sum/length(pcf3_all)
  pcf3_avg <- pcf3_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  
  assign(pcf3, pcf3_avg)
  
  
  # MICROGLIA PCF
  pcf4 <- paste(i, "pcf_mcg", sep="_")

  assign(pcf4,
       foreach(id = 1:length(df)) %dopar% pcf_mcg(df[[id]]))

  pcf4_all <- get(pcf4)
  pcf4_sum <- head(pcf4_all[[1]],n=200)
  for (j in 2:length(pcf4_all)){
    pcf4_sum <- pcf4_sum + head(pcf4_all[[j]],n=200)
  }
  pcf4_avg <- pcf4_sum/length(pcf4_all)
  pcf4_avg <- pcf4_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf4, pcf4_avg)

  
  # -------------------------------GRAPHING-------------------------------
  
  plotto <- paste(i,"plot_pcf",sep="_")
  assign(plotto, 
         ggplot() +
    # B Cells and TAMs
    geom_line(data = pcf1_avg, aes(x=radius, y=mean), color = "yellow4", lwd = 1.5) +
    geom_ribbon(data = pcf1_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "yellow", color = "transparent") +
      
    # B Cells and Tumor Cells
    geom_line(data = pcf2_avg, aes(x=radius, y=mean), color = "cyan4", lwd = 1.5) +
    geom_ribbon(data = pcf2_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "cyan", color = "transparent") +
  
    # B Cells and T Cells
    geom_line(data = pcf3_avg, aes(x=radius, y=mean), color = "red4", lwd = 1.5) +
    geom_ribbon(data = pcf3_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "red", color = "transparent") +
  
    # B Cells and Microglia
    geom_line(data = pcf4_avg, aes(x=radius, y=mean), color = "green4", lwd = 1.5) +
    geom_ribbon(data = pcf4_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "green", color = "transparent") +
      
    labs(x = "Radius (" * mu~"m)", y = "Cell Density", color = "white") +
    coord_cartesian(xlim = c(0, 30), ylim = c(0,600)) +
    theme_classic())
}

stopImplicitCluster()
```

## PCF Analysis B
This analysis will contain all samples. We use B cells as focus point, comparing to TGFB+ TAMs, tumor cells, t cells, and microglia.
```{r pcf_b}
# Load packages
library(doParallel)
library(parallel)
library(foreach)

# Make sure this is the right functions file
source(file = "./source_code/02_functions_b.R")

no_cores <- detectCores()
registerDoParallel(makeCluster(no_cores))


for (i in groups) {
  df <- paste(i, "df", sep = "_")
  
  ls <- conditions %>%
    filter(group == i) %>%
    select_('sample_id')
  
  assign(df,
         foreach(id = 1:nrow(ls)) %dopar% read_mpx(ls, id))

  df <- get(df)

  # TAM PCF
  pcf1 <- paste(i, "pcf_tam", sep="_")

  assign(pcf1,
       foreach(id = 1:length(df)) %dopar% pcf_tam(df[[id]]))
  
  pcf1_all <- get(pcf1)
  pcf1_sum <- head(pcf1_all[[1]],n=200)
  for (j in 2:length(pcf1_all)){
    pcf1_sum <- pcf1_sum + head(pcf1_all[[j]],n=200)
  }
  pcf1_avg <- pcf1_sum/length(pcf1_all)
  pcf1_avg <- pcf1_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf1, pcf1_avg)
  
  
  # TUMOR PCF
  pcf2 <- paste(i, "pcf_tumor", sep="_")

  assign(pcf2,
       foreach(id = 1:length(df)) %dopar% pcf_tumor(df[[id]]))
  
  pcf2_all <- get(pcf2)
  pcf2_sum <- head(pcf2_all[[1]],n=200)
  for (j in 2:length(pcf2_all)){
    pcf2_sum <- pcf2_sum + head(pcf2_all[[j]],n=200)
  }
  pcf2_avg <- pcf2_sum/length(pcf2_all)
  pcf2_avg <- pcf2_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf2, pcf2_avg)

  # T_CELL PCF
  pcf3 <- paste(i, "pcf_tcell", sep="_")

  assign(pcf3,
       foreach(id = 1:length(df)) %dopar% pcf_tcell(df[[id]]))

  pcf3_all <- get(pcf3)
  pcf3_sum <- head(pcf3_all[[1]],n=200)
  for (j in 2:length(pcf3_all)){
    pcf3_sum <- pcf3_sum + head(pcf3_all[[j]],n=200)
  }
  pcf3_avg <- pcf3_sum/length(pcf3_all)
  pcf3_avg <- pcf3_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  
  assign(pcf3, pcf3_avg)
  
  
  # MICROGLIA PCF
  pcf4 <- paste(i, "pcf_mcg", sep="_")

  assign(pcf4,
       foreach(id = 1:length(df)) %dopar% pcf_mcg(df[[id]]))

  pcf4_all <- get(pcf4)
  pcf4_sum <- head(pcf4_all[[1]],n=200)
  for (j in 2:length(pcf4_all)){
    pcf4_sum <- pcf4_sum + head(pcf4_all[[j]],n=200)
  }
  pcf4_avg <- pcf4_sum/length(pcf4_all)
  pcf4_avg <- pcf4_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf4, pcf4_avg)

  
  # -------------------------------GRAPHING-------------------------------
  
  plotto <- paste(i,"plot_pcf",sep="_")
  assign(plotto, 
         ggplot() +
    # B Cells and TAMs
    geom_line(data = pcf1_avg, aes(x=radius, y=mean), color = "yellow4", lwd = 1.5) +
    geom_ribbon(data = pcf1_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "yellow", color = "transparent") +
      
    # B Cells and Tumor Cells
    geom_line(data = pcf2_avg, aes(x=radius, y=mean), color = "cyan4", lwd = 1.5) +
    geom_ribbon(data = pcf2_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "cyan", color = "transparent") +
  
    # B Cells and T Cells
    geom_line(data = pcf3_avg, aes(x=radius, y=mean), color = "red4", lwd = 1.5) +
    geom_ribbon(data = pcf3_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "red", color = "transparent") +
  
    # B Cells and Microglia
    geom_line(data = pcf4_avg, aes(x=radius, y=mean), color = "green4", lwd = 1.5) +
    geom_ribbon(data = pcf4_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "green", color = "transparent") +
      
    labs(x = "Radius (" * mu~"m)", y = "Cell Density", color = "white") +
    coord_cartesian(xlim = c(0, 30), ylim = c(0,600)) +
    theme_classic())
}

stopImplicitCluster()
```

## PCF Analysis C
This analysis will contain all samples. We use B cells as focus point, comparing to TGFB+ TAMs and tumor cells.
```{r pcf_c}
# Load packages
library(doParallel)
library(parallel)
library(foreach)

# Make sure this is the right functions file
source(file = "./source_code/02_functions_b.R")

no_cores <- detectCores()
registerDoParallel(makeCluster(no_cores))


for (i in groups) {
  df <- paste(i, "df", sep = "_")
  
  ls <- conditions %>%
    filter(group == i) %>%
    select_('sample_id')
  
  assign(df,
         foreach(id = 1:nrow(ls)) %dopar% read_mpx(ls, id))

  df <- get(df)

  # TAM PCF
  pcf1 <- paste(i, "pcf_tam", sep="_")

  assign(pcf1,
       foreach(id = 1:length(df)) %dopar% pcf_tam(df[[id]]))
  
  pcf1_all <- get(pcf1)
  pcf1_sum <- head(pcf1_all[[1]],n=200)
  for (j in 2:length(pcf1_all)){
    pcf1_sum <- pcf1_sum + head(pcf1_all[[j]],n=200)
  }
  pcf1_avg <- pcf1_sum/length(pcf1_all)
  pcf1_avg <- pcf1_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf1, pcf1_avg)
  
  
  # TUMOR PCF
  pcf2 <- paste(i, "pcf_tumor", sep="_")

  assign(pcf2,
       foreach(id = 1:length(df)) %dopar% pcf_tumor(df[[id]]))
  
  pcf2_all <- get(pcf2)
  pcf2_sum <- head(pcf2_all[[1]],n=200)
  for (j in 2:length(pcf2_all)){
    pcf2_sum <- pcf2_sum + head(pcf2_all[[j]],n=200)
  }
  pcf2_avg <- pcf2_sum/length(pcf2_all)
  pcf2_avg <- pcf2_avg %>%
    mutate(
      minc = mean - sd,
      maxc = mean + sd)
  assign(pcf2, pcf2_avg)

  
  # -------------------------------GRAPHING-------------------------------
  
  tmp_plot <- paste(i,"plot_pcf",sep="_")
  assign(tmp_plot, 
         ggplot() +
    # B Cells and TAMs
    geom_line(data = pcf1_avg, aes(x=radius, y=mean), color = "yellow4", lwd = 1.5) +
    geom_ribbon(data = pcf1_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "yellow", color = "transparent") +
      
    # B Cells and Tumor Cells
    geom_line(data = pcf2_avg, aes(x=radius, y=mean), color = "cyan4", lwd = 1.5) +
    geom_ribbon(data = pcf2_avg, aes(x=radius, y=mean, ymin = minc, ymax = maxc), alpha = 0.5, fill = "cyan", color = "transparent") +
  
    labs(x = "Radius (" * mu~"m)", y = "Cell Density", color = "white") +
    coord_cartesian(xlim = c(0, 30), ylim = c(0,600)) +
    theme_classic())
}

stopImplicitCluster()
```
